Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-11-09 paper/introduction.tex
Partial updates to introduction

\begin{itemize}
\item
Setting: Consider a producer that creates a data-structure and passes it to a consumer to process. Suppose that the data-structure is very large, and it takes some time to produce it. A GC that occurs during this phase will [potentially, depending on generations] scan the partially constructed data-structure, but this scan is useless since the entire data-structure is alive. In a system, where most of the work takes this form, with producers producing really large data-structures, the GC overheads become significant and better performance can be achieved using alternative memory management techniques. (We could present the Ci\# version of one of motivating Naiad examples here. E.g., it could be the Select example. But when I look at this example, it lacks one of the complicating aspects, namely that a dynamic region may be created in one call of the OnReceive method, updated in several subsequent invocations of the OnReceive method, and finally be transferred in some invocation.)

\item
The goal of this work is to enable transferable regions: essentially a data-structure such as the above. It is best to consider a transferable region as representing a single data-structure (that is, a composite object), and not a collection of unrelated individual objects.

>>>>

Consider the example shown in Fig.~\ref{fig:motivating-eg}.
\textbf{TODO: Replace by more complex example where data-structure is updated
over multiple invocations of method and finally transferred.}
This code represents the logic for a streaming query operator (also known
as a continuous query). The operator receives a continuous stream of
input messages, which are batched into time-windows associated with a logical
timestamp. Messages may be delivered out-of-order and multiple messages with
the same timestamp may be delivered. A timing-message, periodically indicates that
no more input messages with a timestamp less than a given value $t$ will be subsequently
delivered. At this point, the operator can create the output message corresponding
to time $t$ and deliver it to its successors.

\input{motivating-eg}

This example is an instance of a general pattern, where a producer creates a data-structure
and passes it to a consumer.
In a system where most of the computation takes this form and these data-structures are
very large, as is the case with many streaming big-data analysis systems,
the garbage collection overheads become significant.
% as the collector scans these large data-structures
However, these data-structures have a well-defined lifetime, which can be exploited to
achieve better performance 
% using alternative memory management techniques.

Region-based memory management is a good fit for such programs.
A region is a block of memory (consisting of potentially more than one contiguous
range of memory locations, permitting a region to grow) that is used to realize
a data-structure with a well-defined lifetime. In particular, the entire region
is typically freed in one shot.

Region-based memory management, both manual as well as automatic, has been known for
a long time. Manual region-based memory management suffers from the usual challenges,
namely the potential for dangling references and lack of memory safety. Automatic
region-based memory managements guarantee memory safety, but impose various restrictions.
Tofte and Talpin's system, for example, requires that the set of all regions (existing
at a point in time) forms a stack. Specifically, this requires that the lifetimes of
all regions must be well-nested: it is not possible to have two regions whose lifetimes
overlap, with neither one's lifetime contained within the other.
Unfortunately, the data-structures in the above example do not satisfy this restriction.
We refer to such regions as \emph{dynamic} regions.

The goal of this work is an automatic (memory-safe) region-based memory management technique
that supports dynamic regions and higher-order functions.

\begin{itemize}
